Re-examining cross-cultural similarity judgments using language statistics

### 2023-02-02 code/4_paper/manuscript/manuscript.tex
mcf pass

\title{Re-examining cross-cultural similarity judgments using lexical
co-occurrence}

@@

\begin{abstract}
Is ``cow'' more closely related to ``grass'' or ``chicken''? Speakers of
different languages judge similarity in this context differently, but
why? One possibility is that cultures co-varying with these languages
induce differences in conceptualizations of similarity. Specifically,
East Asian cultures may promote reasoning about thematic similarity, by
which cow and grass are more related, whereas Western cultures may bias
judgments toward taxonomic relations, like cow-chicken. This difference
in notions of similarity is the consensus interpretation for
cross-cultural variation in this paradigm. We consider, and provide
evidence for, an alternative possibility, by which notions of similarity
are similar across contexts, but the statistics of the environment vary.
On this account, similarity judgments are guided by co-occurrence in
experience, and hearing about cows and grass or cows and chickens more
often could induce preferences for these groupings, and account in part
for apparent differences in notions of similarity across contexts.

@@

Alternatively, these judgments could be shaped by cross-cultural
differences in the \emph{input} to similarity judgment, that is, the
statistics of the environment, and the content of everyday experiences.
Perhaps when faced with the triad task, participants from all cultures
follow the same process for conceptualizing similarity, but rely on
language or culture-specific input to this process. If we observe a
difference in categorization between East Asian and Western
participants, it could be that members of both groups use the same
procedure (considering similarity that is influenced by both taxonomic
and thematic relations), but the input to this procedure differ between
cultures, with East Asian participants exposed to more support for
thematic similarity in the language they hear in comparison to their
Western counterparts. We might also expect that both conceptualization
of similarity and input from experience play a role in driving
cross-cultural differences -- East Asian participants may be exposed to
more instances of thematic similarity, and prefer to conceptualize
similarity as thematic to a greater extent than Westerners.

@@

To determine whether varied input to similarity judgments can in part
explain cross-cultural differences, we need a way to operationalize
variation in exposure to thematic and taxonomic relations. While
exposure to types of similarity is difficult to measure, the statistics
of language can provide a rough proxy. Language statistics are useful in
that they are part of the input to everyday experience -- and indeed,
may afford many of the `experiences' that people have with infrequently
encountered items, like cows or helicopters -- and they provide an
accessible measure. Previous work also suggests that language
statistics, such as lexical co-occurrence or cosine distance of word
embeddings, can be good predictors of similarity reasoning. Semantic
models that are constructed using lexical co-occurrence (in comparison
to annotated relations) have been shown to perform well on predicting
human judgments about similarity between word pairs that are
thematically or taxonomically related (Rohde, Gonnerman, \& Plaut,
2006). Relatedly, a model trained on word-document co-occurrence can
predict word association and the effects of semantic association on a
variety of linguistic tasks (Griffiths, Steyvers, \& Tenenbaum, 2007).
Word embeddings like word2vec, gloVe, and fastText have also been shown
to be good predictors for similarity judgments Jatnika, Bijaksana, \&
Suryani (2019).

Our study uses cosine distances of fastText word vectors as a measure of
lexical co-occurrence\footnote{We also carried out our analysis using
  raw lexical co-occurrences and obtained similar results.}. fastText is
a system that uses lexical co-occurrence information to generate a
vector representing each word in its lexicon (Mikolov, Grave,
Bojanowski, Puhrsch, \& Joulin, 2018). fastText has also been shown to
be sensitive to cultural differences in word meanings: Thompson \&
Lupyan (2020) demonstrated that the distribution of semantic meaning
clusters generated by fastText trained on language-specific corpora
correlates with the cultural, historical, and geographical similarities
of these languages.

We note that language statistics may incorporate both cultural-specific
environmental statistics, that is, the experiential \emph{input} to
similarity judgments, and culture-specific senses of similarity, the
\emph{conceptualization} of similarity. For example, cultural features
(like farming) can lead to differences in environmental statistics
(seeing cows and grass) and these can influence language (talking more
about cows and grass). But other cultural features (like conceiving of
similarity thematically) could also cause individuals to talk
differently about the same experiences (mentioning what cows eat rather
than what other animals cows are like). Acoordingly, our approach
examines the extent to which language statistics can predict
cross-cultural differences in similarity judgments with the
understanding that language statistics are likely a proxy for both
\emph{input} to and \emph{conceptualization} of similarity.

@@

The present study tests whether the statistics of language can be used
to predict cross-cultural differences in similarity judgments, and
particularly, whether this approach provides additional insight beyond
cross-cultural characterizations based on taxonomic vs thematic
preferences.

@@

This study is correlational and cannot evaluate causal relationships
between environmental statistics and similarity judgments. However, this
work can inform potential mechanisms by examining whether similarity
judgments covary with environmental statistics that differ across these
contexts. Our specific research questions are as follows:~

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\item
  Do we replicate cross-cultural differences in similarity judgments
  between East Asian and Western cultures?
\item
  Are these cross-cultural differences related to differences in
  language?
\item
  Is our language model specific to items with a taxonomic-thematic
  contrast, or can it predict similarity more generally?~
\end{enumerate}

To preview our results, we replicate US-China differences and find that
Vietnamese judgments are intermediate between these two. We find that
language-specific statistics provide good predictions for cross-cultural
differences in similarity judgments, and additionally, for more general
similarity judgments that do not contrast taxonomic and thematic
matches. Critically, the stimuli used to assess this more general case
of similarity reasoning were constructed to be outside the scope of
explanation for taxonomic-thematic accounts (for use as filler items).
While the language model may succeed in taxonomic-thematic similarity
predictions by picking up on differences in the \emph{conceptualization}
of similarity that are reflected in language data, this
conceptualization account provides no prediction for these filler
stimuli. Accordingly, these filler stimuli provide the strongest test of
the language model, demonstrating that it can explain variation in
similarity judgments beyond that explained by thematic or taxonomic
conceptualizations of similarity. Taken together, our findings provide
support for an alternative to previous accounts, on which differing
cultures induce differing conceptions of similarity. They show that
language can account for culture-specific variation in similarity
reasoning, both for taxonomic-thematic judgments and for judgments not
attributable to these senses of similarity. While these findings do not
rule out contributions from differing senses of similarity, they show it
is possible to explain cross-cultural differences in similarity
judgments without invoking variation in notions of similarity, at least
in some cases.

@@

We recruited 200 participants from the US, 199 participants from
Vietnam, and 200 participants from mainland China. US participants were
recruited through snowball sampling seeded with Stanford student email
lists, Vietnam participants were recruited through snowball sampling
seeded with Vietnam-based student groups on Facebook, and mainland China
participants were recruited through snowball sampling seeded with group
chats on WeChat. US participants were compensated with \$5 gift
certificates (USD), VN participants received 50,000₫ (VND) in phone
credit, and mainland China participants received 25CNY through WeChat
credit transfer.

We excluded 8 US participants, 62 Vietnam participants, and 16 China
participants who missed 2 or more attention checks. We followed 4
exclusion criteria that aim to retain only participants who are
influenced by one culture: (1) non-native speakers of English and
Vietnamese, respectively, (2) fluent in at least one of the other two
study languages (Vietnamese for US participants, English for Vietnamese
participants and Chinese participants), (3) have lived outside of the
test country (US, Vietnam, or China) for more than two years, and (4)
have significant international experience (more than 6 international
experiences of 2 days or longer.) We did not use a particular criterion
for a language if it would exclude 25\% or more of any one sample. In
this round of exclusion, we excluded 73 US, 27 Vietnam participants, and
35 China participants. After these exclusions, the US sample included
119 participants (30M, 84F, 3 non-binary, 2 other), with mean age = 22.2
(SD = 8.15) and median age = 20. The Vietnam sample included 110
participants (34M, 71F, 5 other), with mean age = 22.21 (SD = 5.81) and
median age = 21. The China sample included 149 participants (61M, 87F, 1
other), with mean age = 23.1 (SD = 3.65) and median age = 23.\footnote{A
  table summarizing number of participants lost at each round of
  exclusions is included in the Supplementary Information.}

We preregistered a more stringent exclusion criterion where participants
were excluded if they missed any attention checks. However, this led to
a small sample size, especially for Vietnam context (US = 109, China =
132, Vietnam = 57). Our reported results with the less stringent
exclusion criterion (detailed above) is largely not different from the
results with the preregistered criterion. Any differences are noted in
the Analysis section.

\hypertarget{stimuli-materials}{%
\subsection{Stimuli Materials}\label{stimuli-materials}}

We adapted stimuli from previous studies to create a set of test triads
consisting of a cue, with one thematic and one taxonomic match option.
For example, ``cow,'' ``grass,'' and ``chicken,'' where ``cow'' is the
cue, ``grass'' is the thematic match, and ``chicken'' the taxonomic
match. We included 105 such triads, a superset including triads pulled
from supplemental information and in-text examples across the
literature, and others that we adapted or created . We selected triads
on the basis of cultural familiarity in the US, Vietnam, and China. The
triads were originally in English; they were translated to Vietnamese
and Mandarin by a fluent bilingual speaker in each language. The
translations were then checked for accuracy after backtranslation to
English by another fluent bilingual in each language who was naive to
the original English versions. All materials are available at {[}BLINDED
FOR REVIEW{]}.

@@

Each participant completed all 105 triads in sets of 21 trials at a time
(10 test triads, 10 filler triads, and 1 attention check per page), by
selecting the match most related to the cue (``Which thing is most
closely related to the bolded item?'' and translated equivalents). The
test triads were presented with 105 filler triads mixed in, to obscure
the taxonomic-thematic two-answer forced choice structure of the test
stimuli and reduce the likelihood that participants would become aware
of the design. The filler triads were groups of three semantically
related words, but where the match options were not distinguished by
thematic vs.~taxonomic similarity, e.g., cue ``bird'' with match options
``lizard'' and ``toad.'' Additionally, we included 10 attention check
trials, which were formatted like the test and filler triads but
included an instruction instead of a cue item, e.g., ``Choose wife''
with match options ``wife'' and ``husband.'' In total, each participant
completed 210 similarity judgments and 10 attention check questions,
with triads presented in randomized orders that varied between subjects.

@@

We use the fastText pre-trained models of English, Mandarin, and
Vietnamese in Grave, Bojanowski, Gupta, Joulin, \& Mikolov (2018). These
models are trained on Common Crawl and Wikipedia using We distribute
pre-trained word vectors for 157 languages, trained on Common Crawl and
Wikipedia using fastText. These models were trained using a Continuous
Bag of Words (CBOW) with position-weights and a window of size 5. The
models use character n-grams of length 5 and 10 negative examples. From
the aforementioned models, we retrieve the word vectors (dimension 300)
for each word we are interested in.

@@

To give an intuition for our model, consider again the cow-grass-chicken
triad: we retrieved word vectors for ``cow'' and ``grass'', and
calculate the cosine distance between these vectors. Similarly, we
retrieved vectors for ``cow'' and ``chicken'' and calculate the cosine
distance between them. Our similarity prediction is then inversely
proportional to the ratio of cosine distance of these pairs. This is
because a larger cosine distance means the word vectors are further
apart, and thus the words are less similar. For example, if the cosine
distance of thematic cow-grass is 0.7 and the cosine distance of
taxonomic cow-chicken is 0.3, then our model predicts, correspondingly,
that 30\% of responses to the triad will be grass, and the other 70\%
chicken.

In practice, we calculated the cosine distance between each cue-thematic
match (thematic cosine distance) and cue-taxonomic match (taxonomic
cosine distance), using the spatial.distance.cosine function from the
SciPy package (Virtanen et al., 2020). We then calculated the thematic
cosine distance proportion as thematic cosine distance over the sum of
taxonomic cosine distance and thematic cosine distance. We did this for
all three corpora. We were able to obtain predictions for all triads in
all languages. We then use a mixed-effects regression to evaluate how
well each corpus model predicts participants' similarity judgments,
across triads and cultural contexts. All analyses are available at
{[}BLINDED FOR REVIEW{]}.

@@

Following previous work, we would expect participants from mainland
China to prefer thematic matches more than US participants (to use the
cow-grass-chicken triad: we would expect participants from mainland
China to prefer the cow-grass match over the cow-chicken match to a
larger extent compared to US participants). We would also expect that
participants from Vietnam would pattern with China, and also show a
significantly higher preference for thematic matches compare to US
participants.

The group means of proportion of thematic response in mainland China is
the highest (M = 0.65, SD = 0.11), followed by the groups means in
Vietnam (M = 0.6, SD = 0.11), which is slightly higher than that of the
US (M = 0.56, SD = 0.17) (Figure 1).

@@

To test for cross-context differences in similarity judgments between
the countries, we ran a mixed-effects logistic regression predicting
triad responding (taxonomic or thematic) with country (US, China, or
Vietnam) as a fixed effect. As random effects, we included an intercept
per subject and one per triad, as well as by-triad random slopes for
country to account for variation in the country effect across triads.

Overall, there is a significant effect of country on proportion of
thematic responses (\(\chi^2\)(2) = 15.37, p \textless{} .001). However,
this effect is driven by the difference between US and China responding
(\(\beta\) = -0.48, p \textless{} .001). There is no statistical
difference between the Vietnam and China responding (\(\beta\) = -0.22,
p = 0.09), and the US and Vietnam responding (\(\beta\) = 0.26, p =
0.086).

On this analysis, we do not find support that the US-China tendencies
toward taxonomic and thematic responding (respectively) extend to the
US-Vietnam comparison. Accordingly, we cannot speak to overall biases
toward thematic responding across Asian cultural contexts broadly, but
we do replicate the differences documented by Ji et al. (2004) between
the US and China. However, in our corpus model comparison, we do find
evidence for different, more fine-grained variation in similarity
judgments between the US and Vietnam.

@@

To test whether variation in language statistics can explain differences
in similarity judgments between US and Vietnam participants, we compare
logistic mixed-effects regression models fit to the thematic responding
data from each country separately. We first ask how well each corpus
model (English, Vietnamese, or Mandarin) predicts similarity judgments
by speakers of the corresponding language (US, Vietnam, or China). To do
this, we use a mixed-effects logistic regression to predict triad
responses (0=taxonomic or 1=thematic) with corpus prediction (proportion
of cosine distance) as a fixed effect and participant and triad as
random effects. If environmental statistics (as proxied by language
statistics) contribute to the differences in similarity judgments, we
would expect each language corpus to be a good predictor for similarity
judgment responding in its corresponding context.

We found that all corpora are significant predictors of all cultural
context responding, with p \textless{} 0.05 and \(\beta\) from -8.69 to
-2.4. (For a full report, see Supplementary Information.)

While each corpus is a good predictor for its corresponding context, the
fact that all corpora covary with all cultural contexts suggests that
language is such a rich proxy of human experience that even using the
wrong proxy (e.g.~Mandarin for US responding) is informative. A single
corpus model might therefore not be informative in culture-specific
ways, but might only reflect consistency in experiences across cultures.

\hypertarget{multiple-corpora-model}{%
\subsubsection{Multiple corpora model}\label{multiple-corpora-model}}

If language statistics is able to predict meaningful culture-specific
variation in similarity judgment (rather than just consistency across
cultures), we would expect each corpus to be the best predictor of its
corresponding culture compared to the other two corpora. We directly
compare the corpus models by including both as fixed effects in three
mixed-effect regressions (predicting US, Vietnam and China responding)
with the same random effects as above.

For US responding: only the English (EN) corpus is a significant
predictor\footnote{With preregistered exclusion criterion: only the
  English (EN) and Mandarin (ZH) corpus are significant predictors.}. EN
corpus: \(\beta\) = -6.96, \(\chi^2\)(1) = 16.57, p \textless{} .001. VI
corpus: \(\beta\) = -2.22, \(\chi^2\)(1) = 3.73, p = 0.054. ZH corpus:
\(\beta\) = -3.18, \(\chi^2\)(1) = 3.37, p = 0.066.

For Vietnam responding: only the Vietnamese (VI) and Mandarin (ZH)
corpus are significant predictors\footnote{With preregistered exclusion
  criterion: all three corpora are significant predictors.}. EN corpus:
\(\beta\) = -2.98, \(\chi^2\)(1) = 2.58, p = 0.108. VI corpus: \(\beta\)
= -2.75, \(\chi^2\)(1) = 4.84, p = 0.028. ZH corpus: \(\beta\) = -4.39,
\(\chi^2\)(1) = 5.49, p = 0.019.

For China responding: only the Mandarin (ZH) and English (EN) corpus are
significant predictors. EN corpus: \(\beta\) = -3.32, \(\chi^2\)(1) =
7.18, p = 0.007. VI corpus: \(\beta\) = -1.59, \(\chi^2\)(1) = 3.63, p =
0.057. ZH corpus: \(\beta\) = -5.31, \(\chi^2\)(1) = 17.69, p
\textless{} .001.

We observed some level of language specificity from this analysis. The
English corpus is the best predictor for US responding, and the Mandarin
corpus is the best predictor for China response. While this is not the
case with the Vietnamese corpus and the Vietnam responding, the
Vietnamese corpus is still a significant predictor for the Vietnam
responding (Figure 2). These results strengthens our hypothesis that
variation in inputs to similarity judgment (as proxied by language
statistics) can predict cross-cultural variations of similarity
judgment.

@@

However, in all cultural contexts, adding the other two corpora produces
a significantly better fit than the identical model without the
additional corpora, and only the corresponding corpus included as a
predictor (US response: \(\chi^2\)(2) = 7.93, p = 0.019; Vietnam
response: \(\chi^2\)(2) = 13.72, p = 0.001; China response:
\(\chi^2\)(2) = 10.56, p = 0.005). This analysis suggests that
culture-specific inputs to similarity judgment (as proxied by language
statistics) do not fully explain cross-cultural differences in
similarity judgment.

@@

A possible concern with our approach is that our models using language
statistics in Q2 are only able to account for variation in
thematic/taxonomic preference in similarity judgment because language
statistics is picking up on differences in thematic/taxonomic preference
in conceptualization of similarity. For example, perhaps ``cow'' and
``grass'' co-occurs more than ``cow'' and ``chicken'' in Mandarin
because Chinese speakers prefer thematic relations, and thus talk about
``cow'' and ``grass'' together more often. Vice versa, English speakers
preferring taxonomic relations would talk about ``cow'' and ``chicken''
together more often, hence driving higher lexical co-occurrence for
``cow'' and ``chicken'' in the English corpus. In other words, perhaps
the language statistics models are merely measuring variation in
thematic/taxonomic relation preference in conceptualization of
similarity, only in a different way. If this is the case, language
statistics should not be able to predict responding for filler items
(such as whether ``tomorrow'' or ``yesterday'' is more related to
``today'') because these items do not have a thematic/taxonomic
structure. Additionally, even when one option might be more related to
the cue, such relationship is not systematic throughout the set of
fillers. On the other hand, if language statistics capture similarity
judgment in a more general sense (beyond thematic/taxonomic preference
in conceptualization of similarity), it should also be a significant
predictor for the non-structured filler items.

>>>>

\title{Re-examining cross-cultural similarity judgments using language
statistics}

@@

\begin{abstract}
Is ``cow'' more closely related to ``grass'' or to ``chicken''? Speakers
of different languages judge similarity in this context differently, but
why? One possibility is that cultures co-varying with these languages
induce differences in conceptualizations of similarity. Specifically,
East Asian cultures may promote reasoning about thematic similarity, by
which cow and grass are more related, whereas Western cultures may bias
judgments toward taxonomic relations, like cow and chicken. We measure
similarity judgments across the US, China, and Vietnam and replicate
US-China cultural differences with Vietnam intermediate. We also show
that word embedding models (fastText models for each language) are
related to judgments within each country, suggesting a possible
alternative interpretation of cross cultural differences. Perhaps
notions of similarity are similar across contexts, but the statistics of
the linguistic environment vary.

@@

Alternatively, these judgments could be shaped by cross-cultural
differences in the \emph{input} to similarity judgment, that is, the
statistics of the linguistic and/or physical environment as manifest in
everyday experiences. Perhaps when faced with the triad task,
participants from all cultures follow the same process for
conceptualizing similarity, but rely on language or culture-specific
input to this process. If we observe a difference in categorization
between East Asian and Western participants, it could be that members of
both groups use the same procedure (considering similarity that is
influenced by both taxonomic and thematic relations), but that the input
to this procedure differ between cultures, with East Asian participants
exposed to more support for thematic similarity in their experience in
comparison to their Western counterparts.

@@

To determine whether varied input to similarity judgments can in part
explain cross-cultural differences, we need a way to operationalize
variation in exposure to thematic and taxonomic relations. While
exposure to types of similarity in everyday experience is difficult to
measure, the statistics of language can provide a rough proxy. Language
statistics are useful in that they are part of the input to everyday
experience -- and indeed, may afford many of the ``experiences'' that
people have with infrequently encountered items, like cows or
helicopters -- and they provide an accessible measure. Previous work
also suggests that language statistics, such as lexical co-occurrence or
cosine distance of word embeddings, can be good predictors of similarity
reasoning. Semantic models that are constructed using lexical
co-occurrence (in comparison to annotated relations) have been shown to
perform well on predicting human judgments about similarity between word
pairs that are thematically or taxonomically related (Rohde, Gonnerman,
\& Plaut, 2006). Relatedly, a model trained on word-document
co-occurrence can predict word association and the effects of semantic
association on a variety of linguistic tasks (Griffiths, Steyvers, \&
Tenenbaum, 2007). Word embeddings like word2vec, gloVe, and fastText
have also been shown to be good predictors for similarity judgments Asr,
Zinkov, \& Jones (2018).

Our study uses cosine distances of fastText word vectors as a measure of
lexical co-occurrence.\footnote{We also carried out our analysis using
  raw lexical co-occurrences and obtained similar results.} fastText is
a system that uses lexical co-occurrence information to generate a
vector representing each word in its lexicon (Mikolov, Grave,
Bojanowski, Puhrsch, \& Joulin, 2018). Practically speaking, fastText
vectors are available for many languages, and fastText has also been
shown to be sensitive to cultural differences in word meanings: Thompson
\& Lupyan (2020) demonstrated that the distribution of semantic meaning
clusters generated by fastText trained on language-specific corpora
correlates with the cultural, historical, and geographical similarities
of these languages.

Language statistics may incorporate both cultural-specific environmental
statistics, that is, the experiential \emph{input} to similarity
judgments, and culture-specific senses of similarity, the
\emph{conceptualization} of similarity. For example, cultural features
(like farming) can lead to differences in environmental statistics
(seeing cows and grass) and these can influence language (talking more
about cows and grass). But other cultural features (like conceiving of
similarity thematically) could also cause individuals to talk
differently about the same experiences (mentioning what cows eat rather
than what other animals cows are like). Accordingly, our approach
examines the extent to which language statistics can predict
cross-cultural differences in similarity judgments with the
understanding that language statistics are likely a proxy for both
\emph{input} to and \emph{conceptualization} of similarity.

@@

@@

This study is correlational and cannot evaluate causal relationships
between environmental statistics and similarity judgments. However, this
work can inform potential mechanisms by examining whether similarity
judgments co-vary with environmental statistics that differ across these
contexts.

Our specific research questions are as follows:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\item
  Do we replicate cross-cultural differences in similarity judgments
  between East Asian and Western cultures?
\item
  Are these cross-cultural differences related to differences in
  language?
\item
  Is our language model specific to items with a taxonomic-thematic
  contrast, or can it predict similarity more generally?
\end{enumerate}

To preview our results, we replicate US-China differences and find that
Vietnamese judgments are intermediate between these two. We find that
language-specific statistics provide good predictions for cross-cultural
differences in similarity judgments, and additionally, for more general
similarity judgments that do not contrast taxonomic and thematic
matches. Critically, the stimuli used to assess this more general case
of similarity reasoning were constructed to be outside the scope of
explanation for taxonomic-thematic accounts (for use as filler items).
While the language model may succeed in taxonomic-thematic similarity
predictions by picking up on differences in the \emph{conceptualization}
of similarity that are reflected in language data, this
conceptualization account provides no prediction for these filler
stimuli.

Taken together, our findings provide support for an alternative to
previous accounts on which differing cultures induce differing
conceptions of similarity. On this alternative, language explains for
culture-specific variation in similarity reasoning, both for
taxonomic-thematic judgments and for judgments not attributable to these
senses of similarity. While these findings do not rule out other
cross-cultural factors, they show it is possible to explain some
cross-cultural variation in similarity judgments purely from language
statistics.

@@

Data collection and analyses for this study were pre-registered, and the
pre-registration is available at {[}blinded for peer review{]}. We
recruited 200 participants from the US, 199 from Vietnam, and 200 from
mainland China. All participants were recruited through snowball
sampling, which, in the US was seeded with student email lists at a
large university, in Vietnam (VN) with Vietnam-based student groups on
Facebook, and in China (CN) with student groups on WeChat. US
participants were compensated with \$5 gift certificates (USD), VN
participants received 50,000₫ (VND) in phone credit, and CN participants
received 25CNY through WeChat credit transfer.

We excluded 8 US participants, 62 VN participants, and 16 CN
participants who missed two or more of the 10 attention check questions.

In doing so, we deviated from our preregistered exclusion criterion,
which used a more stringent policy, excluding participants who missed
any of the 10 attention checks. This approach would have substantially
reduced our sample size, disproportionately affecting the VN sample
(resulting in a final sample of 109 US, 132 CN, and 57 VN participants).
We report results with the less stringent exclusion criterion, but we
performed analyses with both samples and found broadly similar results.
Any differences are noted in the Analysis.

We also applied four demographic exclusion criteria designed to reduce
cross-cultural influences across our samples: we excluded participants
who (1) were not native speakers of the test language (English,
Vietnamese, or Mandarin Chinese), (2) were fluent in another one of the
study languages, (3) had lived outside of the test country for more than
two years, or (4) had significant international experience (more than 6
international experiences of 2 days or longer). Following our
pre-registration, we dropped any exclusion criterion that would exclude
25\% or more of any one population, but only for that population.
Specifically, we did not exclude VN and CN participants who spoke
English, or US participants with significant international experience.
As a result of these demographic exclusions, we excluded 73 US, 27 VN,
and 35 CN participants. The final US sample included 119 participants
(30M, 84F, 3 non-binary, 2 other), with mean age = 22.2 (SD = 8.15). Our
VN sample included 110 participants (34M, 71F, 5 other), with mean age =
22.21 (SD = 5.81). And the CN sample included 149 participants (61M,
87F, 1 other), with mean age = 23.1 (SD = 3.65).{[}\^{}4{]}

\hypertarget{materials}{%
\subsection{Materials}\label{materials}}

All stimuli and scripts for the experiment and analysis are available at
{[}BLINDED FOR REVIEW{]}. We collected and adapted stimuli from previous
studies to create test triads consisting of a cue with one thematic and
one taxonomic match option. For example, ``cow,'' ``grass,'' and
``chicken,'' where ``cow'' is the cue, ``grass'' is the thematic match,
and ``chicken'' the taxonomic match. We included 105 such triads, a
superset including triads pulled from supplemental information and
in-text examples across the literature, and others that we adapted or
created . We selected triads on the basis of cultural familiarity within
the US, Vietnam, and China. The triads were translated from English to
Vietnamese and Mandarin by fluent bilingual speakers of each language,
and back-translated to English by another fluent bilingual who was naive
to the original English version. The back-translations were checked
against the original English to identify ambiguities in the translation,
which were resolved through discussion and selection of alternative
terms when relevant.

@@

Each participant completed all 105 triads in blocks of 21 trials at a
time (10 test triads, 10 filler triads, and 1 attention check per page),
by selecting the match most related to the cue (``Which thing is most
closely related to the bolded item?'' and translated equivalents). The
test triads were presented with 105 filler triads mixed in, to obscure
the taxonomic-thematic two-answer forced choice structure of the test
stimuli and reduce the likelihood that participants would become aware
of the design. The filler triads were groups of three semantically
related words, but where the match options were not distinguished by
thematic vs.~taxonomic similarity, for example, the cue ``bird'' with
match options ``lizard'' and ``toad.'' Additionally, we included 10
attention check trials, which were formatted similarly to the test and
filler triads but included an instruction instead of a cue item, e.g.,
``Choose wife'' with match options ``wife'' and ``husband.'' In total,
each participant completed 210 similarity judgments (105 test triads and
105 fillers) and 10 attention check questions, with all items presented
in randomized orders that varied between subjects.

@@

We use the fastText pre-trained models of English, Mandarin, and
Vietnamese in Grave, Bojanowski, Gupta, Joulin, \& Mikolov (2018). These
models were trained on Common Crawl and Wikipedia using fastText and use
character n-grams of length 5, and 10 negative examples. The training
used a Continuous Bag of Words with position-weights and a window of
size 5. From these models, we retrieve the word vectors (dimension 300)
for each word in our triad and filler stimuli.

@@

To give an intuition for our model, consider again the cow:
grass/chicken triad. We retrieved word vectors for ``cow'' and
``grass'', and calculated the cosine distance between these vectors. We
did the same for ``cow'' and ``chicken.'' Our similarity prediction is
inversely proportional to the ratio of cosine distance of these pairs.
This is because a larger cosine distance means the word vectors are
further apart, and thus the words are less similar. For example, if the
cosine distance of thematic cow-grass is 0.7 and the cosine distance of
taxonomic cow-chicken is 0.3 (the more similar of the two), then our
model predicts, correspondingly, that 30\% of responses to the triad
will be grass, and the other 70\% chicken.

We calculated the cosine distance between each cue-thematic match
(thematic cosine distance) and cue-taxonomic match (taxonomic cosine
distance). We then calculated the thematic cosine distance proportion as
thematic cosine distance over the sum of taxonomic cosine distance and
thematic cosine distance. We followed this process for all three
corpora, and were able to obtain predictions for all triads in all
languages. Then, we used a mixed-effects regression to evaluate how well
each corpus model predicts participants' similarity judgments, across
triads and cultural contexts.

@@

In keeping with Ji et al. (2004), we would expect participants from
mainland China to prefer thematic matches more than US participants
(e.g., to prefer the cow-grass match over cow-chicken to a greater
extent than US participants). On the basis of the previous literate more
broadly, we would also expect participants from Vietnam to pattern with
China, showing a stronger preference for thematic matches than US
participants.

We observed the strongest preference for thematic matching in China (M =
0.65, SD = 0.11), followed by Vietnam (M = 0.6, SD = 0.11), and then the
US (M = 0.56, SD = 0.17; see Figure 1).

@@

To test for differences in similarity judgment between the countries, we
ran a mixed-effects logistic regression predicting triad match
(taxonomic or thematic) with country (US, China, or Vietnam) as a fixed
effect. As random effects, we included an intercept per subject and one
per triad, as well as by-triad random slopes for country to account for
variation in the country effect across triads.

We found a significant effect of country (\(\chi^2\)(2) = 15.37, p
\textless{} .001), but this effect is driven by the difference between
US and CN responding (\(\beta\) = -0.48, p \textless{} .001). There was
no statistical difference between matching in Vietnam and China
(\(\beta\) = -0.22, p = 0.09), or the US and Vietnam (\(\beta\) = 0.26,
p = 0.086).

In sum, we replicated the differences documented by Ji et al. (2004)
between the US and China. We did not find that Vietnamese participants
grouped particularly with Chinese participants, however. Instead they
were intermediate between the two. Accordingly, we did not find strong
support for the hypothesis that there are overall biases toward thematic
responding across Asian cultural contexts broadly.

@@

First, to test whether variation in language statistics can explain
differences in similarity judgments, we use a mixed-effects logistic
regressions fit individually for each country. Our models predicted
responses by individual participants to particular triads (0=taxonomic
or 1=thematic) with fastText predictions (proportion of cosine distance)
as a fixed effect and participant and triad as random effects. If
language statistics, contribute to the differences in similarity
judgments, we would expect each language corpus to be a good predictor
for similarity judgment responding in its corresponding context.

All corpora were significant predictors of all cultural context
responding, with p \textless{} 0.05 and \(\beta\) from -8.69 to -2.4.

\hypertarget{multi-corpus-model}{%
\subsubsection{Multi-corpus model}\label{multi-corpus-model}}

If language statistics are able to predict meaningful culture-specific
variation in similarity judgments (rather than just consistency across
cultures), we would expect each corpus to be the best predictor of its
corresponding culture compared to the other two corpora. We directly
compared the corpus models by including all three corpus predictors as
fixed effects in three mixed-effect regressions (predicting US, Vietnam
and China responding) with the same random effects as above.

For US responding: only the English (EN) corpus was a significant
predictor\footnote{With the preregistered exclusion criterion, only the
  English (EN) and Mandarin (ZH) corpus were significant predictors.}.
EN corpus: \(\beta\) = -6.96, \(\chi^2\)(1) = 16.57, p \textless{} .001.
VI corpus: \(\beta\) = -2.22, \(\chi^2\)(1) = 3.73, p = 0.054. ZH
corpus: \(\beta\) = -3.18, \(\chi^2\)(1) = 3.37, p = 0.066.

For Vietnam responding, only the Vietnamese (VI) and Mandarin (ZH)
corpus were significant predictors\footnote{With the preregistered
  exclusion criterion, all three corpora were significant predictors.}.
EN corpus: \(\beta\) = -2.98, \(\chi^2\)(1) = 2.58, p = 0.108. VI
corpus: \(\beta\) = -2.75, \(\chi^2\)(1) = 4.84, p = 0.028. ZH corpus:
\(\beta\) = -4.39, \(\chi^2\)(1) = 5.49, p = 0.019.

For China responding, only the Mandarin (ZH) and English (EN) corpus
were significant predictors. EN corpus: \(\beta\) = -3.32, \(\chi^2\)(1)
= 7.18, p = 0.007. VI corpus: \(\beta\) = -1.59, \(\chi^2\)(1) = 3.63, p
= 0.057. ZH corpus: \(\beta\) = -5.31, \(\chi^2\)(1) = 17.69, p
\textless{} .001.

We observed some language specificity in this analysis (Figure 2). The
English corpus was the best predictor for US responding, and the
Mandarin corpus was the best predictor for Chinese responding. While
this is not the case with the Vietnamese corpus and the Vietnam
responding, the Vietnamese corpus was still a significant predictor for
the Vietnam responding. These results support our hypothesis that
specific language statistics can predict cross-cultural variations of
similarity judgment.

@@

However, in all cultural contexts, adding the other two corpora produced
a significantly better fit than the identical model with only the
corresponding corpus included as a predictor (US response: \(\chi^2\)(2)
= 7.93, p = 0.019; Vietnam response: \(\chi^2\)(2) = 13.72, p = 0.001;
China response: \(\chi^2\)(2) = 10.56, p = 0.005). This analysis
suggests that culture-specific inputs to similarity judgment (as proxied
by language statistics) do not fully explain cross-cultural differences
in similarity judgment, whether due to noise in the estimates of
language statistics or true additional variance picked up by embeddings
from other corpora.

@@

One possible concern with our approach is that our models are only able
to account for variation in similarity judgment because the word
embedding models pick up on cross-cultural differences in the
conceptualization of similarity. Perhaps some aspect of the geometry of
the embedding space specifically reflects these differences rather than
capturing more item-specific patterns of co-occurrence.

If this is the case, the embedding models should be less accurate at
predicting responding for filler items (such as whether ``tomorrow'' or
``yesterday'' is more related to ``today'') because these items do not
have a thematic/taxonomic structure. Additionally, even when one option
might be more related to the cue, that relationship is not systematic
throughout the set of fillers. On the other hand, if the embedding
models capture similarity judgments more generally (beyond
thematic/taxonomic preferences), they should predict judgments for the
unstructured filler items as well as they do for our experimental items.

For each filler triad, we randomly assigned one of the responding
options as `Word1'. Running a mixed-effects logistic regression to
predict responding (Word1 or Word2) with country as a fixed effect and a
random effect structure as above Q1, we found no effect of country on
filler responding (\(\chi^2\)(2) = 0.74, p = 0.692).

Using the same mixed-effects logistic regression structure as the single
corpus models in Q2, we predicted responses (1=word1 or 0=word2) in each
cultural context with the corresponding corpus predictor as a fixed
effect, and participant and triad as random effects. In all cultural
contexts, the corresponding corpus was a significant predictor of
responding, with p \textless{} 0.05 and \(\beta\) from -9.59 to -3.14.
These results show that the word embedding models predict general
similarity in addition to structured thematic/taxonomic similarity.
